---
title: "Class 11: Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Working with the online Protein Data Bank (PDB)

We moved the csv files from PDB to our document to allow analysis of protein structure. 
After downloading the data export summary, we are going to look at the composition of this database. 

```{r}
stats <- read.csv("Data Export Summary.csv", row.names = 1)
stats
```

Percentage of structures solved by x-ray and EM.
```{r}
pre.by.method <- stats$Total/sum(stats$Total) * 100
names(pre.by.method) <- rownames(stats)
pre.by.method
```

What portion of the structures are proteins?
```{r}
round(sum(stats$Proteins)/sum(stats$Total) * 100, 2)
```

How do we bring in data from a file that is NOT .csv? We need a specific package. It's called "r datapasta". Then we can use the "Addins" tab and use the "paste as" function.

```{r}
tmp <- data.frame(stringsAsFactors=FALSE,
   Experimental.Method = c("X-Ray", "NMR", "Electron Microscopy", "Other",
                           "Multi Method", "Total"),
              Proteins = c(124770, 10988, 2057, 250, 127, 138192),
         Nucleic.Acids = c(1993, 1273, 31, 4, 5, 3306),
    ProteinNA.Complex = c(6451, 257, 723, 6, 2, 7439),
                 Other = c(10, 8, 0, 13, 1, 32),
                 Total = c(133224, 12526, 2811, 273, 135, 148969)
)
View(tmp)
```

##Using VMD
We're going to be looking at the structure of 1HSG (HIV-1 Protease)
(All in class notes).
In this case- we searched 1HSG and opened the .pdb file in VMD, which allowed us to visualize the molecular structure. 

##Getting Started with the Bio3D Package

Make sure we have the package open in this session (since it is already downloaded).
```{r}
library(bio3d)
```

Open the 1HSG protein information that we have. 
```{r}
pdb <- read.pdb("1hsg")
pdb
```


Let's install a new package (this one is not quite up to date yet). Allows us to use programs that are from papers- latest, and is very helpful. 

The 'devtools' package allows us to install development versions
  install.packages("devtools")
Install the bio3d.view package from bitbucket
  devtools::install_bitbucket("Grantlab/bio3d-view")

```{r}
library("bio3d.view")
view(pdb, "overview", col="sse")
```

We want to extract the protein only portion of this PDB structure and write it to a new PDB file. 

Extract the ligand (in this instance , the drug) alone. 

##Atom Selection in Bio3D

We'll be using the atom.select() function. This function operates on PDB structure objects (as created by read.pdb()) and returns the numeric indices of a selected atom subset. These indices can then be used to access the $atom and $xyz attributes of PDB structure related objects.

For example to select the indices for all C-alpha atoms we can use the following command:
(Any protein, ligand, or other protein factor can be used in place of "calpha")
```{r}
ca.inds <- atom.select(pdb, "calpha")
ca.inds
```

Let's print the details from the first few selected atoms. 
```{r}
head( pdb$atom[ca.inds$atom, ] )
```
Printing the selected x,y,z coordinates
```{r}
head( pdb$xyz[, ca.inds$xyz] )
```

Quick example of calling the ligand. 
```{r}
lig.inds <- atom.select(pdb, "ligand")
lig.inds
```

```{r}
lig.inds$atom
pdb$atom[lig.inds$atom,]
```

Make a new structure object. 
```{r}
ligand.pdb <- trim.pdb(pdb, lig.inds)
ligand.pdb
```

```{r}
view(ligand.pdb)
```

```{r}
write.pdb(ligand.pdb, file = "ligand.pdb")
```

We skipped to the ligands, but let's do the same thing with the protein (that we called the alpha carbons for).
```{r}
ca.pdb <- trim.pdb(pdb, ca.inds)
ca.pdb

write.pdb(ca.pdb, file = "ca.pdb")
```

Now let's write this out for the entire protein. 
*Add this in later*
(instead of = ligand, use = protein)

##Visualizing a Dynamic Enzyme
```{r}
library(bio3d)
```


Now we're going to look at movement in an enzyme (this one is the egg white enzyme).
```{r}
pdb2 <- read.pdb("1HEL")

#Normal Mode Analysis
modes <- nma(pdb2)
plot(modes)
```

```{r}
m7 <- mktrj(modes, mode=7, file = "mode_7.pdb")
view(m7, col=vec2color( rmsf(m7) ))
```

We used the Bio3D package, and it is important to remember this package is still being updated.
