---
title: "Class 12: Structural Bioinformatics and Drug Discovery"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cleaning up the Protein Target Structure

Start by downloading the target (the protein receptor) from the main PDB database. This time we'll be using the PDB ID "1hsg" the HIV protease. 

Here we're writing a code that will allow us to switch out the protein (why we have the pdb.code defined).

```{r}
library(bio3d)

pdb.code <- "1hsg"
file.name <- get.pdb(pdb.code)
```

This file is very large, so we want to extract the protein only segments. We'll then change it to a new PDB format file. We'll also be doing this for the bound ligand (will allow us to manipulate both).

```{r}
hiv <- read.pdb(file.name)
hiv
```

Initially, we'll want just the protein. When you run it, check the non protein values (should be zero).
```{r}
prot <- trim.pdb(hiv, "protein")
prot
```

We give the file a generic name to avoid hardocding 1hsg (keeps our code versatile).
```{r}
prot.filename <- paste(pdb.code, "_protein.pdb", sep = "")
write.pdb(prot, file = prot.filename)
```

Doing the same for the ligand.
```{r}
lig <- trim.pdb(hiv, "ligand")
lig
```
```{r}
lig.filename <- paste(pdb.code, "_ligand.pdb", sep = "")
write.pdb(lig, file = lig.filename)
```

##Using The AutoDock Tools Package

Once we open the 1hsg_protein.pdb file, we add back the hydrogens and save the file as a .pdbqt
Opening it we can see more of the electrostatic charges. Adding the hydrogens back gives us a lot of room to play around. It also assigns the charges to each molecule. 

We can do the exact same thing for the ligands. 

##Converting Docking Results for Viewing in VMD

```{r}
res <- read.pdb("all.pdbqt", multi = TRUE)
res
```

```{r}
write.pdb(res, file = "results.pdb")
```

We can also compare docking results in R using RMSD (Root Mean Square Distance)
```{r}
#We already ran the first part of this function.
#res <- read.pdb("all.pdbqt", multi=TRUE)
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```

This will compare all of the pairs (14x14 table).
```{r}
rmsd(res)
```

##Ligand Based Approaches- Normal Mode Analysis







