---
title: "Class 6 R Functions"
author: "Katelynn Kazane"
date: "1/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## File reading (again!)

Here we try to use **read.table()** and friends to input some example data into R. 

Let's insert a code chunk.
```{r}
read.table("https://bioboot.github.io/bggn213_S18/class-material/test1.txt", header = TRUE, sep = ",")
```

```{r}
file1 <- "https://bioboot.github.io/bggn213_S18/class-material/test1.txt"
data1 <- read.csv(file1)
data1
```

```{r}
read.table("https://bioboot.github.io/bggn213_S18/class-material/test2.txt", header = TRUE, sep = "$")
```
```{r}
file2 <- "https://bioboot.github.io/bggn213_S18/class-material/test2.txt"
data2 <- read.table(file2, header = TRUE, sep = "$")
data2
```

```{r}
read.table("https://bioboot.github.io/bggn213_S18/class-material/test3.txt")
```

```{r}
file3 <- ("https://bioboot.github.io/bggn213_S18/class-material/test3.txt")
data3 <- read.table(file3)
data3
```

## R functions
My first function! 

```{r}
add <- function(x, y=1) {
# Sum the input x and y
x + y
}
```

Let's use the **add()** function 
```{r}
add(1)
```

```{r}
add(1,5)
```

You can even add to vectors
```{r}
add( c(1,2,3,4))
```

```{r}
add(c(1,2,3,4), 4)
```

If you type {r eval=FALSE} then the program will skip over a chunk of code (important if you're knitting a document). In the following case, we're showing code that will break the function add() so we can either set eval=FALSE or turn the code into a comment. 
```{r}
#Need to make 1,2,2 a vector since our function can only add two numbers. The sum function already exists.
#add(1, 2, 2)

add(1, y=2)
sum(1,2,2)
```
```{r}
add(c(1,2,2))
```

```{r}
#Can't add a numeric and a string. 
#add(x=1, y="b")
```

## Our 2nd Function

```{r}
rescale <- function(x) {
rng <-range(x)
(x - rng[1]) / (rng[2] - rng[1])
}
```

Here we test our function (pick something simple that you know the answer to).
```{r}
rescale(1:10)
```


```{r}
# How would you get your function to work here…
rescale( c(1,2,NA,3,10) )
```
```{r}
x <- rescale( c(1,2,NA,3,10) )
rng <- range(x)
rng

```

```{r}
rescale2 <- function(x) {
rng <-range(x, na.rm = TRUE)
(x - rng[1]) / (rng[2] - rng[1])
}
```
```{r}
rescale2( c(1,2,NA,3,10) )
```

```{r eval=FALSE}
rescale( c(1,2,3,10, "bggn213") )

```

```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
  if(na.rm) {
  rng <-range(x, na.rm=na.rm)
  } else {
  rng <-range(x)
  }
print("Hello")
answer <- (x - rng[1]) / (rng[2] - rng[1])
print("is it me you are looking for?")
if(plot) {
plot(answer, typ="b", lwd=4)
}
print("I can see it in ...")
}
```

```{r}
rescale3(1:10, plot = TRUE)
```

##Using the Bio3D Package

```{r}
library(bio3d)
```

```{r}
pdb <- read.pdb("1hbs")
```

```{r}
pdb
```

```{r}
# Can you improve this analysis code?
library(bio3d)

s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

##Making the Biopackage Example into a Function

```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

