---
title: "Class 7 Functions and Packages"
author: "Katelynn Kazane"
date: "1/30/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Functions revisited

Load the rescale() function from *class 6*. 

```{r}
source("http://tinyurl.com/rescale-R")
```

Now we're going to test this function. 

```{r}
rescale(1:5)
```

```{r}
#rescale(c(1:5, "string"))
```

We want to make the code more robust to these types of errors- meaning mixing strings and numerics. 

```{r}
#rescale2(c(1:5, "string"))
```

Adding these modifications can make understanding the code easier for users. More concise error messages are always better. 

```{r}
is.numeric(1:5)
```
```{r}
is.numeric("string")
```
```{r}
is.numeric(c(1:5, "string"))
```

```{r}
!is.numeric( c(1:5, "string"))
```
```{r}
!is.numeric(1:5)
```

```{r}
# Lets define an example x and y
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
is.na(x)
is.na(y)
```

Now we're looking for what position is TRUE (rather than having to read through everything individually).

```{r}
is.na(x)
is.na(y)

is.na(x) & is.na(y)
```
We want to know how many TRUE's are in the last line. 

```{r}
is.na(x)
is.na(y)

is.na(x) & is.na(y)
sum(is.na(x) & is.na(y))
# which(is.na(x) & is.na(y)) will tell us where the true values are. 
```

We have this working snippet, and will make it into a function. 

```{r}
both_na <- function(x, y) {
  ## Check for NA elements in both input vectors 
  sum( is.na(x) & is.na(y) )
}
```

```{r}
both_na(x,y)
```

Testing this function more! 

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
y3 <- c(1, NA, NA, NA, NA)
```

```{r}
#What will this return?

both_na(x,y2)
```
This is a little weird, because the shorter function is repeating. 

```{r}
both_na(x, y3)
```
We can help resolve this error by adding an IF function. 

```{r}
both_na(x, y2)
```

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

both_na(x,y)
```

